Randomly-Scoped Lambda Calculus

### 2013-01-19 RSLC/paper/paper.tex
Polish up rslc paper

Consider the following lambda calculus term:
\[ (\lambda x. (\lambda x. x))~A~B \]
In ordinary lambda calculus, this always evaluates to $B$ (quite unforgivingly so, in my opinion).
However, I give the programmer the benefit of the doubt, and assume they named the first argument $x$ as well because they wanted it to have equal opportunity. Hence, in RSLC this term can evaluate to either $A$ or $B$.
(On the other side of the coin, RSLC encourages good programming practice: a programmer who {\em doesn't} want $x$ to evaluate to $A$ should name the first argument something different, to make their intentions more clear.)

@@

Now that I have shown in an abstract sense how RSLC can be used to express the same deterministic computations from conventional lambda calculus as well as some novel random programming techniques, let's do a real-world application.
Using the uniform randomness combinator from Section~\ref{sec:uniform}, we can implement any number of standard dice for any nonzero number of sides:

\begin{eqnarray*}
	\mathsf{d}(\Suc{N}) &=& \mathsf{add}~1~\mathsf{rand}[0,N] \\
       2\mathsf{d}(\Suc{N}) &=& \mathsf{let}~x=\mathsf{add}~\mathsf{rand}[0,N]~\mathsf{rand}[0,N] \\
	&& \mathsf{in}~\mathsf{add}~2~x \\
       3\mathsf{d}(\Suc{N}) &=& \mathsf{let}~x=\mathsf{add}~\mathsf{rand}[0,N]~\mathsf{rand}[0,N] \\
	&& \mathsf{in}~\mathsf{let}~y=\mathsf{add}~x~\mathsf{rand}[0,N] \\
	&& \mathsf{in}~\mathsf{add}~3~y
\end{eqnarray*}
and so on.

@@

>>>>

Consider the following lambda calculus term:
\[ (\lambda x. (\lambda x. x))~A~B \]
In ordinary lambda calculus, this always evaluates to $B$ (quite unforgivingly so, in my opinion).
However, I give the programmer the benefit of the doubt, and assume they named the first argument $x$ as well because they wanted it to have equal opportunity. Hence, in RSLC this term can evaluate to either $A$ or $B$.
(On the other side of the coin~\cite{twoface}, RSLC encourages good programming practice: a programmer who {\em doesn't} want $x$ to evaluate to $A$ should name the first argument something different, to make their intentions more clear.)

@@

Now that I showed in an abstract sense how RSLC can be used to express popular computations from conventional lambda calculus as well as some novel random programming techniques, let's do a real-world application.
Using the uniform combinator from Section~\ref{sec:uniform}, we can make standard dice for any nonzero number of sides:

\begin{eqnarray*}
	\mathsf{d}(\Suc{N}) &=& \mathsf{add}~1~\mathsf{rand}[0,N] \\
       2\mathsf{d}(\Suc{N}) &=& \mathsf{let}~x=\mathsf{add}~\mathsf{rand}[0,N]~\mathsf{rand}[0,N] \\
	&& \mathsf{in}~\mathsf{add}~2~x \\
       3\mathsf{d}(\Suc{N}) &=& \mathsf{let}~x=\mathsf{add}~\mathsf{rand}[0,N]~\mathsf{rand}[0,N] \\
	&& \mathsf{in}~\mathsf{let}~y=\mathsf{add}~x~\mathsf{rand}[0,N] \\
	&& \mathsf{in}~\mathsf{add}~3~y
\end{eqnarray*}

@@

My implementation of RSLC in Haskell is available at \url{https://github.com/bblum/sigbovik/blob/master/RSLC/RSLC.hs}.
